Space is the Place: Effects of Continuous Spatial Structure on Analysis of Population Genetic Data

### 2019-10-11 ms/review-responses.tex

\reply{
Good suggestion -- the SFS in Fig 3A is now on a log-log scale, which shows the slight decrease in low frequency SNPs a little better. 
}

\begin{point}{Figure S3:}
    Caption seems to be missing detail 
\end{point}

\reply{
    Thanks for catching this - we have revised this caption to add details including the simulation parameters. \llname{ll:figs3caption}
}

%%%%%%%%%%%%%%
\reviewersection{1}
We thank the reviewer for their very constructive comments. Responses follow below:

@@

\begin{point}{}
    When neighbourhood size is small (10-100), the mean number of IBS tracts > 2bp (nIBS as in Table S1) is elevated similar to Wright's inbreeding coefficient, but mean of the distribution of pairwise IBS (mean(IBS)) is decreased. What could be the source of this discrepancy? How exactly mean(IBS) was calculated?
\end{point}

\reply{
The mean of the IBS tract distribution is calculated by building a (very large) vector of lengths of all IBS tracts between all pairs of individuals and taking the mean over this entire vector. This distribution is quite strongly skewed (i.e. see figure 4 and the skew(IBs) panel of fig S1) such that the highest density is in low values. We explain the depressed mean value by noting that at low neighborhood sizes the coalescence time of distant (in space) individuals is inflated, which allows more time for recombination and mutation to break up IBS tracts. The flip side of this is that there are then more IBS tracts in total across a random sample of individuals, explaining the corresponding inflation in the number of tracts. We also see an increase in the number of very long IBS tracts at low neighborhood sizes, which corresponds to an excess of very recent coalescent events for groups of nearby individuals. 
\cjb{need to think of something to say we did about this in the text}

@@

\begin{point}{\revref}
    In the sentence, 'We also examined p values for systemic inflation' I think the authors meant 'systematic inflation'.
\end{point}

\reply{
    Whoops; thanks. Fixed.
}

\begin{point}{}
    Page 11, Please correct the legend in Figure 2: must be 'spatial model' and 'random mating' model.
\end{point}

\reply{
Thanks for catching our confusing legend title placement! We have moved "model" to after "spatial" as suggested.
}

@@

    \begin{enumerate}
        \item It nice to actually see spatially explicit simulations and I'm happy that forward simulation is now fast enough that you can do this sort of thing.
        \item The paper is very clear and well-written, easy to understand the motivation and most of the details. That's not always the case for this sort of paper.
        \item I felt that the section about the effect on GWAS was the most interesting and novel part of the paper and gave me some intuition that I hadn't had before.
    \end{enumerate}

    I don't have any major criticisms. There were a few aspects that I thought might warrant some additional discussion, and a few specific questions below. The general questions I had after reading it were:
\end{quote}

\begin{point}{}
    To what extent are any of the results dependent on the exact method of simulation. There are a number of choices about the exact details of the simulations (e.g. the way the overlapping generations are handled, the edge effects and, particularly, the form of Equation 1 - see below). It's not so much that these are non-standard (since I don't think there is a standard) and they all sort of make sense heuristically, and I was left wondering whether these sorts of choices actually make a difference. Do the authors have some thoughts/intuition/results about that? Given that the results in Fig. 3 seem quite consistent with expectations, I suspect that on some level it doesn't make much difference but then there are intermediate results like Fig. 2 which seem a bit counter-intuitive and I wonder if those aspects depend on the simulation scheme.
\end{point}

\reply{
This is a good point also raised by other reviewers. We have added an appendix section comparing our model to a reverse-time stepping stone simulation on several relevant summary statistics \llname{appendix1}. We show that for $\theta_\pi$ and several other statistics the stepping stone model approaches the continuous model as the resolution of the landscape increases. We also see some interesting differences that seem to reflect artifacts from discretization. We are also curious how all the other analyses we test would be affected using other simulation schemes, but hope to explore that aspect in future work. 
}

@@

\begin{point}{}
    Related to the first point, to what extent are the results qualitatively different to those that would be obtained in a stepping-stone model? My interpretation is that they are actually very similar, but I didn't see whether that was explicitly discussed. In some sense, it's still easier to do large simulations in a stepping-stone model so it would be nice to be reassured that that's still ok.
\end{point}

\reply{
    See above. It is certainly easier and faster, but comes with significant issues either when neighborhood size is lower than the population of a deme, or when demes are small enough that the sample size approaches local $N_e$, at least in reverse-time simulations. 
}

@@

\begin{point}{}
    The authors use a scaling factor in equation (2) to counteract the increase in fitness of individuals at the edges. Can they provide a figure showing that this is the case. What does "roughly" mean on line 164. Perhaps a heatmap of the fitness of individuals across the grid with and without the scaling factor?
\end{point}

\reply{
Good suggestion -- we have added a supplemental figure to show the distribution of individual fitness across the landscape with and without our edge-scaling approach. MAKE THIS FIG AND POINT TO IT HERE.
}

\begin{point}{}
    It would be helpful provide the figure showing that generating mutations during the forward simulations in SLIM is equivalent to applying mutations using msprime on pre-generated trees (line 185)? It sounds like this procedure would underestimate the variance in the number of mutations, since you remove the effect of random generation time. Is this effect small?
\end{point}

\reply{
    We have added a figure showing sample site-frequency-spectra generated from a subset of simulations run with SLiM mutations, and then using msprime to apply mutations to the same tree sequences with our generation-time scaling approach \ref{fig:slim_v_msp_sfs}. These approaches yield extremely similar spectra so we believe any error in the approximation is quite small. We also now show variation in simulation times with different mutation and tree sequence recording schemes in Figure \ref{fig:runtimes}, and included a brief discussion of this in \llname{runtimesxn}
}

### 2019-11-09 ms/spaceness.tex
removed redundant figure versions (replaced with symbolic links)

\begin{figure}[htbp]
\centering
\includegraphics{ms/figures/sampling_maps.pdf}
\caption{Example sampling maps for 60 individuals on a 50$\times$50 landscape for midpoint, point, and random sampling strategies, respectively.}
\label{fig:samplemap}
\end{figure}

@@

\begin{figure*}[p]
\centering
\includegraphics[width=\textwidth]{ms/figures/sfs_w_sumstats_loglog.pdf}
\caption{Log-scaled site frequency spectrum (A) and summary statistic distributions (B) by sampling strategy and neighborhood size.}
\label{fig:sumstats}
\end{figure*}

@@

\begin{figure*}[p]
\centering
\includegraphics[width=.5\textwidth]{ms/figures/ibs_cumdists_w_rm.pdf}
\caption{Cumulative distributions for IBS tract lengths per pair of individuals at different geographic distances, across three neighborhood sizes (NS).}
\label{fig:ibs}
\end{figure*}

@@

\begin{figure*}[h]
\centering
\includegraphics[width=\textwidth]{ms/figures/grid_ss_curves_edit.pdf}
\caption{Summary statistics for 2-dimensional coalescent stepping-stone models with fixed total $N_{e}$ and varying numbers of demes per side. The black "infinite" points are from our forward-time continuous space model. Inter-deme migration rates are related to $\sigma$ as described above.}
\label{fig:grid_stats}
\end{figure*}

@@

\begin{figure*}[p]
\centering
\includegraphics[width=\textwidth]{ms/figures/wm_v_dd_maps.pdf}
\caption{Maps of individual locations in a continuous-space Wright-Malécot model with independent dispersal of all individuals (top) and under our continuous space model incorporating density-dependant fitness (bottom). The clustering seen in the top row is the "Pain in the Torus" described by \citet{Felsenstein1975}.}
\label{fig:slim_v_msp_sfs} 
\end{figure*}
\FloatBarrier

>>>>

\begin{figure}[htbp]
\centering
\includegraphics{figures/sampling_maps.pdf}
\caption{Example sampling maps for 60 individuals on a 50$\times$50 landscape for midpoint, point, and random sampling strategies, respectively.}
\label{fig:samplemap}
\end{figure}

@@

\begin{figure*}[p]
\centering
\includegraphics[width=\textwidth]{figures/sfs_w_sumstats_loglog.pdf}
\caption{Log-scaled site frequency spectrum (A) and summary statistic distributions (B) by sampling strategy and neighborhood size.}
\label{fig:sumstats}
\end{figure*}

@@

\begin{figure*}[p]
\centering
\includegraphics[width=.5\textwidth]{figures/ibs_cumdists_w_rm.pdf}
\caption{Cumulative distributions for IBS tract lengths per pair of individuals at different geographic distances, across three neighborhood sizes (NS).}
\label{fig:ibs}
\end{figure*}

@@

\begin{figure*}[h]
\centering
\includegraphics[width=\textwidth]{figures/grid_ss_curves_edit.pdf}
\caption{Summary statistics for 2-dimensional coalescent stepping-stone models with fixed total $N_{e}$ and varying numbers of demes per side. The black "infinite" points are from our forward-time continuous space model. Inter-deme migration rates are related to $\sigma$ as described above.}
\label{fig:grid_stats}
\end{figure*}

@@

\begin{figure*}[p]
\centering
\includegraphics[width=\textwidth]{figures/wm_v_dd_maps.pdf}
\caption{Maps of individual locations in a continuous-space Wright-Malécot model with independent dispersal of all individuals (top) and under our continuous space model incorporating density-dependant fitness (bottom). The clustering seen in the top row is the "Pain in the Torus" described by \citet{Felsenstein1975}.}
\label{fig:slim_v_msp_sfs} 
\end{figure*}
\FloatBarrier
